---
title: "Furlong Exam 3"
author: "Morgan Furlong"
date: "'r format(Sys.time(), '%d %B %Y')'"
output: html_notebook
  html document: 
  df_print: paged 
  word_document: default 
  html_motebook: default 
  pdf_document: default 
editor_options: 
  chunk_output_type: inline 
---

```{r docinfo, echo = FALSE}

#get runtime 
date()

#session info 
sessionInfo()

#loading required packages 
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Hmisc)

```

```{r import, echo=FALSE}

#Import data  
crime <- read_csv("~/Downloads/nola_crime_2018.csv")

class(crime)

#dataset on New Orleans crime statistics 2018 

```

Question 1: 
9,970 observations

Question 2: 
24 variables 

```{r, echo=FALSE}

Unique <- unique(crime$Item_Number)
length(Unique)

```
Question 3: 
Out of the 9,970 item numbers, 9,384 are unique 

```{r, echo=FALSE}

crime <- crime[!duplicated(crime$Item_Number), ]

Irish <- crime %>% 
  filter(crime$GNOCDC_LAB == "IRISH CHANNEL") %>% select(Signal_Description) 

sort(Irish$Signal_Description)

```

Question 4: Out of the 43 offense reports described in the Irish Channel, 1 was categorized as "aggravated". 

```{r, echo=FALSE}

neighbor <- group_by(crime, GNOCDC_LAB)
class(neighbor)

neigh <- summarise(neighbor, Item_Number = n()) %>% arrange(Item_Number)
neigh

tail(neigh)

```
Question 5: Little Woods and the French Quarter have the largest number of reported crimes compared to other neighborhoods. 

```{r, echo=FALSE}

population <- crime %>% select(GNOCDC_LAB, pop) %>% group_by(GNOCDC_LAB) %>% summarise(mean_pop = mean(pop, na.rm = TRUE))

population <- population[order(population$GNOCDC_LAB),]
population
neigh <- neigh[order(neigh$GNOCDC_LAB),]
neigh

rate <- ((neigh$Item_Number / population$mean_pop) * 1000) 
rate

```

Question 6: Table 1. Comparing crime abundance and rate per 1000 persons in New Orleans, LA (2018)
```{r table, echo=TRUE}

rate <- cbind(neigh,rate) 
colnames(rate) [1] <- "Neighborhood"
colnames(rate) [2] <- "Number of Crime Reports"
colnames(rate) [3] <- "Rate of Crime per 1000 Persons"
rate


```

```{r, echo=FALSE}

arr_population <- population %>% arrange(mean_pop)
tail(arr_population)

```
Question 7: The largest neighborhood by population in the dataset is Little Woods with a total population of 44,947 persons. 

```{r, echo=FALSE}

arr_rate <- rate %>% arrange(`Rate of Crime per 1000 Persons`)
head(arr_rate)
population %>% filter(GNOCDC_LAB == "LAKEWOOD")

```
Question 8: Lakewood, with a population size of 13,989 and crime report number of 10, had the lowest crime rate of 0.71 per 1000 persons. 


```{r creating value, echo=FALSE}

ratio <- rate$`Rate of Crime per 1000 Persons`

```

```{r rate ratio, echo=TRUE }
Lakewood <- ratio / 0.71
Lakewood

FrenchQuarter <- ratio / 44.16
FrenchQuarter

```
Question 9: Table 2. Rate ratio with Lakewood Population as reference. 

Question 10: Table 3. Rate ratio with French Quarter population as reference. 

```{r, echo=FALSE}

sort(table(crime$Signal_Description),decreasing = TRUE)
sort(table(crime$Charge_Description),decreasing = TRUE)

```

Question 11: Domestic disturbance was the most common cause for a police report to be filed, but the most common charge that was filed is theft. 

```{r, echo=FALSE}

dom <- neighbor %>% select(Signal_Description, GNOCDC_LAB, pop)
dom %>% group_by(GNOCDC_LAB) 
dom %>% summarise(mean_pop = mean(pop, na.rm = TRUE))
domestic <- dom %>% filter(str_detect(Signal_Description,"DOMESTIC")) 
domesticdisturbance <- domestic %>% count(Signal_Description) %>% group_by(GNOCDC_LAB) %>% summarise(n = sum(n))

ddrate <- ((domesticdisturbance$n / population$mean_pop) * 1000) 
ddrate
```

```{r table, echo=TRUE}

cbind(domesticdisturbance$GNOCDC_LAB,ddrate)
```
Question 12: Table 4. The rate of domestic disturbance by neighborhood in New Orleans, LA. 

```{r, echo=FALSE}
#rate of all crimes per neighborhood based on average age of victims 

reg <- neighbor %>% select(Victim_Age,GNOCDC_LAB) %>% group_by(GNOCDC_LAB) %>% summarise(Victim_Age = mean(Victim_Age, na.rm = TRUE))

```

```{r regression, echo=TRUE}

linearmod <- lm(rate$`Rate of Crime per 1000 Persons` ~ reg$Victim_Age)      
print(linearmod)
summary(linearmod)
```
Question 13: Linear Regression Model (Above)
```{r tabulate, echo=FALSE}

hypo <- (rate$`Rate of Crime per 1000 Persons`) / (population$mean_pop)
hypo
hypothesis <- cbind(population$GNOCDC_LAB,hypo) 



```

Question 14: When comparing rates of crime reports to population, Village de Lest may have the worst police responsiveness. 

```{r histogram, echo=TRUE}

pois <- rpois(10000, lambda = 4.7)
hist(pois, main = "Histogram of Randomly Distributed Poisson", xlab = "events (x)")

```
Question 15: Figure 1: Poisson distribution of randomly distributed variable where lambda = 4.7. 

```{r loop, echo=TRUE}

iris
species <- split(iris,iris$Species)
str(species)
res <- list()
for(n in names(species)){
  data <- species[[n]]
  res[[n]] <- data.frame(species=n, mean.sepal.length=mean(data$Sepal.Length))
}  
print(res)
```
Question 16: The average sepal length for the three species of irises are printed above. 

```{r plot, echo=TRUE}

cars <-data.frame(mtcars)
scatter.smooth(cars$mpg,cars$disp)
ggplot(mtcars) +
  geom_point(aes(x = mpg, y = disp, color = cyl)) + labs(title = "Comparison of Car Performance", x = "mpg", y = "displacement (cu.in)") 
```
Question 17: Figure 2: Comparison of car performance based on displacement and number of cylinders and their effect on miles per gallon. (legend = number of cylinders)


```{r vector, echo = TRUE}

x <- c(-4.89, -1.93, -1.11, 3.94, 0.46, -3.85, -0.20, 6.04,9.36, 7.26, 4.88, 13.45, -2.93, 6.39, -16.22, -3.24, 6.86, 11.87, 1.81, -2.54)

u <- mean(x)
u
o <- sd(x)
o

Z <- (x-u)/o
Z
```
Question 18: The Z scores for the specified vector are above. 


