# clear console, global environment  ----
rm(list = ls())
cat("\014")  #sends 'CTRL + L' to console oprional

---
title: "Furlong_Take_Home3"
author: "Morgan Furlong"
date: "11/18/2019"
output: html_document
html_notebook: default
  html_document:
    df_print: paged
---

```{r setup, include = FALSE}
#-------------------------------------------------------------------------
# Take home assigment 4: Mosquito Control in Ghana          
#    M. Furlong 21 November 2019  mfurlon@tulane.edu

#    This script contains the code used to anaylze the effect of insecticide use (applyed to bed nets) on childhood mortality from Malaria in Ghana as part of take home assignment 4

#   Dependencies: Script requires use of the tidyverse and devtools packages. All other function calls are contained in base R installs as of version 3.5.3    
#-------------------------------------------------------------------------
```

#compile date and session info 

```{r docinfo, echo=FALSE}
#run libraries 
library(reprex)
library(tidyverse)

#install.packages("Hmisc") 
library(Hmisc)

#print and get run time 
date()
current_date <- date()

#session info 
devtools::session_info()
sessionInfo()

```

# Loading data 

```{r load, echo=FALSE}
data_df <- read_csv("~/Desktop/Intro to Stats Packages/Statistical Software Tutorial 1/ghana_bednet_malaria.csv")[ ,1:8]

```


# Data Summarization 

```{r summary, echo=FALSE}

summary(data_df)

```

#Question 1: This dataset contains 26,342 observations. 

#Question 2: This dataset contains 8 variables. 

```{r describe, echo=TRUE }

describe(data_df[7])

```

#Question 3: 50.5% of the children represented in this study are male.

```{r sum, echo=TRUE}
describe(data_df$follyr)

follyr_female <- data_df %>% filter(sex == 1)

describe(follyr_female)

sum(follyr_female$follyr, na.rm = TRUE)

```

#Question 4: The total number of years of follow-up experienced by female children in intervention areas was 16,473. 

```{r calculation, echo=TRUE}

outcome_52 <- data_df %>% filter(cluster == 52)

describe(outcome_52)

typeof(outcome_52$outcome)

typeof(outcome_52$sex)

storage.mode(outcome_52$outcome) <- "character"

typeof(outcome_52$outcome)

describe(outcome_52$outcome)

```

#Question 5: 2.4% of children in cluster 52, an intervention cluster, died during follow-up. 

```{r analysis, echo=TRUE}

mean(data_df$agemn)

sd(data_df$agemn, na.rm = FALSE)

```

#Question 6: The average subject age was 25.3 months old (sd = 18.02)


# Calculate mortality rate in control and intervention arm 
```{r}

control_mort <- data_df %>% filter(bednet == 0)
intervention_mort <- data_df %>% filter(bednet == 1)

control_mortrate <- control_mort %>% filter(outcome == 1)
intervention_mortrate <- intervention_mort %>% filter(outcome == 1)

control_mortality_rate <- nrow(control_mortrate) / sum(control_mort$follyr, na.rm = TRUE)

intervention_mortality_rate <- nrow(intervention_mortrate) / sum(intervention_mort$follyr, na.rm = TRUE)


```

# Question 7: The mortality rate of the intervention arm is 2.4% compared to 2.8% in the control arm. 

```{r }

data_df %>% group_by(cluster) %>%
  summarise(prop = sum(outcome, na.rm = TRUE)/n()) %>%
  barplot(prop ~ cluster, data = ., main = "Proportion Died per Cluster")

```
#Question 8: Bar graph showing proportion of subjects who died in each cluster. 

```{r logistic model, echo=TRUE}

logitmod <- glm(data_df$outcome ~ data_df$bednet, family = binomial(link = "logit"))

print(logitmod) 

```
#Question 9: Regression Summary 

#Question 10: The subjects in the intervention cluster demonstrated a 17% reduction in mortality (RR = 0.83; 95% CI: 0.69-1.00; p = 0.05). (Risk ratio is used since the health outcome is common). 



